---
title: "Penguin Flipper Lengths ANOVA"
author: "Teresa Fukuda"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data
Packages: tidyverse, car, onewaytests


```{r, include=FALSE}

library (tidyverse)
library (car)

penguins <- read_csv("penguins.csv")


```
Our question: Is there a significant difference in mean flipper length for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago?

Maybe I can use a one way ANOVA...
- Independent observations
- Equal variances (Levene's test)
- Normality (histograms and qq plots)
Only after exploring these things can we decide whether or not an ANOVA is appropriate to test these means. 



```{r}
# For Gentoo penguins, how many females and males were observed (and for all other species, too)?

#within the dataset penguins, create a table with species and sex
#table() creates a really quick contingency table with values

counts <- with(penguins, table(species, sex)) 

# update the data with less difficult species names

penguins2<- penguins %>%
  mutate(
    common_name =
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter (sex == "FEMALE")


```

