---
title: "Penguin Flipper Lengths ANOVA"
author: "Teresa Fukuda"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data
Packages: tidyverse, car, onewaytests


```{r, include=FALSE}

library (tidyverse)
library (car)

penguins <- read_csv("penguins.csv")


```
Our question: Is there a significant difference in mean flipper length for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago?

Maybe I can use a one way ANOVA...
- Independent observations
- Equal variances (Levene's test)
- Normality (histograms and qq plots)
Only after exploring these things can we decide whether or not an ANOVA is appropriate to test these means. 



```{r}
# For Gentoo penguins, how many females and males were observed (and for all other species, too)?

#within the dataset penguins, create a table with species and sex
#table() creates a really quick contingency table with values

counts <- with(penguins, table(species, sex)) 

# update the data with less difficult species names

penguins2<- penguins %>%
  mutate(
    common_name =
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter (sex == "FEMALE")


```

Exploratory graphs and Levene's test for equal variances

```{r}

hists <- ggplot(penguins2, aes(x=flipper_length))+
  geom_histogram(aes(fill=common_name)) + 
  facet_wrap(~common_name, scales = "free")

hists

# At this point ask, does it look like there might be a significant difference between at least two groups??

qqplots <- ggplot(penguins2, aes(sample=flipper_length)) +
  geom_qq(aes(color=common_name)) +
  facet_wrap(~common_name, scales= "free")

qqplots
# More evidence that the populations seem to be normal anyway, even with small sample sizes

# Based on qq plots and histograms, observations from each group appear normally distributed. Also n>30 for all groups, so means will be normally distributed by Central Limit Theorem. YOU MUST DO THIS IN ORDER TO CONFIRM THAT THE TEST YOU ARE DOING IS APPROPRIATE. 


```

Evaluate variances and perform Levene's test.

```{r}
# If the largest sample variance is less than 4x greater than the smallest sample variance, then these are close enough to consider variances equal (an ANOVA is robust enough for this)

#group_by can take more than one argument

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance=var(flipper_length)
  )

variances
# The largest variance (33.1) is definitely not more than 4x greater than the smallest variance (15.2)

# Levene's test
# leveneTest function in the "car" package

# Takes argument of variable as a function of the factor
levene_penguin <- leveneTest(flipper_length ~ common_name, data= penguins2)
levene_penguin

#H0: variances are equal
#H1: vairances are not equal

# Retain null hypothesis, p>0.05; there is no significant difference in variances across species. 

# With a larger sample size, more likely to have a low p value, but then check actual values of variances 

```

One Way ANOVA by penguin flipper length by species (common_name)

```{r}
# H0: the means across all groups are equal
# H1: the means between at least two groups are different (or all three differ)

penguin_aov <- aov(flipper_length ~ common_name, data= penguins2)
summary(penguin_aov)
#use summary to get actual ANOVA outcome

# Very unlikely to have randomly gotten samples with means this different, if the actual means are the same. 

# Which ones differ??
# Post hoc testing by Tukey's HSD

# H0: no significant difference in pairwise means
# H1: there is a significant difference in pairwise means

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

# There are differences between all of the pairwise means!
# aov.test function in the 'onewaytests' package allows for inline

```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p< 0.001, $\alpha$ = 0.05 with post hoc Tukey's HSD, $\alpha$ = 0.05).
